---
title: "Using the web application"
description: >
  Documentation for using the web application
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the web application}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

See the [Quick Start](../index.html#web-application-quick-start)

## Supported Input Formats

### 1) pMN CSV plate

|                   |               |
| ----------------- | ------------- |
| Description       | CSV input from a plate reader used for the pMN assay |
| Format            | CSV  |
| Extension         | `.csv`  |
| Example           | [`example_data_pmn_plate_n1.csv`](https://github.com/PhilPalmer/AutoPlate/blob/main/data-raw/example_data_pmn_plate_n1.csv) | |
| Required columns  | `WellPosition`, `RLU` |
| Plate number      | The end of the filename can be used to specify the filename eg `1.csv` for plate 1. If numbers are not provided the filename will be used instead |
| Number of files   | One CSV file per plate. Multiple files can be uploaded |


### 2) ELLA Excel plate

|                   |               |
| ----------------- | ------------- |
| Description       | Excel input from a plate reader used for the ELLA assay |
| Format            | Excel  |
| Extension         | `.xls`, `xlsx` |
| Example           |  |
| Required columns  |  |
| Plate number      | The end of the filename can be used to specify the filename eg `1.xls` for plate 1. If numbers are not provided the filename will be used instead |
| Number of files   | One Excel file per plate. Multiple files can be uploaded |


### 3) AutoPlate CSV plates

|                   |               |
| ----------------- | ------------- |
| Description       | CSV input downloaded from AutoPlate on the QC page generated from one of the two previous input files |
| Format            | CSV  |
| Extension         | `.csv`  |
| Example           | [`example_data_pmn_platelist_H1N1.csv`](https://github.com/PhilPalmer/AutoPlate/blob/main/data-raw/example_data_pmn_platelist_H1N1.csv) |
| Required columns  | `types`, `sample_id`, `dilution`, `bleed`, `virus`, `rlu`, `neutralisation`, `treatment`, `experiment_id`, `exclude`, `plate_number` |
| Plate number      | The plate number will be specified in the `plate_number` column |
| Number of files   | One CSV file which can correspond to multiple plates  |

### Columns
```{r, echo=FALSE, layout="l-body-outset"}
column_descriptions_filepath <- "../data-raw/example_data_column_descriptions.csv"
example_data_column_descriptions <- utils::read.csv(column_descriptions_filepath,
  header=TRUE,
  stringsAsFactors=FALSE,
  check.names=FALSE
)
knitr::kable(example_data_column_descriptions)
```