---
title: "Interactive pMN Assay Report - Generated by AutoPlate"
output:
  html_document:
    code_folding: hide
    number_sections: TRUE
params:
  drm_model: NA
---

# Input

## Load the AutoPlate R library
```{r}
# install.packages("devtools")
# devtools::install_github("PhilPalmer/AutoPlate", dependencies = TRUE)
library(autoplate)
```

## Title

```{r}
platelist_file <- "pmn_platelist.csv"
assay_df <- utils::read.csv(platelist_file, header = TRUE, stringsAsFactors = FALSE, check.names = FALSE)
paste(toString(unique(assay_df$experiment_id)), "- Bleed", toString(unique(assay_df$bleed)), "- Virus", toString(unique(assay_df$virus)))
```

## Let's take a look at the data

```{r, layout="l-body-outset"}
knitr::kable(head(assay_df, 3))
```

## Dimensions before exclusion

```{r}
paste0("Rows: ", dim(assay_df)[1], ", Columns: ", dim(assay_df)[2])
```

## Dimensions after exclusion

```{r}
data <- dplyr::filter(assay_df, exclude == FALSE)
paste0("Rows: ", dim(data)[1], ", Columns: ", dim(data)[2])
```

## Number of wells excluded

```{r}
dim(assay_df)[1] - dim(data)[1]
```

# Quality Control

## Average Virus + Cell Luminescence

```{r, layout="l-body-outset"}
av_lum_df <- init_av_lum_df(assay_df)
knitr::kable(av_lum_df)
```

## Heatmaps for The 96 Well Plate(s)

```{r}
features <- c("Types", "Sample ID", "Dilution", "Virus", "RLU", "Neutralisation", "Treatment", "Experiment ID", "Bleed", "Exclude")
plates <- sort(unique(assay_df$plate_number))

for (feature in features) {
  for (plate in plates) {
    plot_heatmap(plate, assay_df, gsub(" ", "_", tolower(feature), fixed = TRUE), feature)
  }
}
```

## Types boxplot

```{r}
# Preprocessing
boxplot_data <- data %>%
  dplyr::mutate(types = ifelse((types == "c"), "cell", types)) %>%
  dplyr::mutate(types = ifelse((types == "m"), "monoclonal antibody", types)) %>%
  dplyr::mutate(types = ifelse((types == "v"), "virus", types)) %>%
  dplyr::mutate(types = ifelse((types == "x"), "serum sample", types))
boxplot_data$plate_number <- as.factor(boxplot_data$plate_number)

# Generate plot
types_boxplot <- init_types_boxplot(boxplot_data)
plotly::ggplotly(types_boxplot) %>% plotly::layout(boxmode = "group")
```

# Results

## Data Exploration

```{r}
# Preprocessing
data <- dplyr::filter(data, types %in% c("x", "m"))

# Generate plot
data_exploration_plot <- plot_data_exploration(data)
plotly::ggplotly(data_exploration_plot)
```

## DRM model

```{r}
drm_model <- params$drm_model
drm_model
```

## Dose Response Curve(s)

```{r}
# Produce a seperate plot for each virus
htmltools::tagList(lapply(unique(data$virus), function(virus_to_plot) {
  # Preprocessing
  data <- dplyr::filter(assay_df, types %in% c("x", "m"), exclude == FALSE, virus == virus_to_plot)
  model <- eval(parse(text = paste("drc::drm(", drm_model, ")")))
  # Generate plot
  drc_plot <- plot_drc(data, model)
  plotly::ggplotly(drc_plot)
}))
```

## Individual Effective Dose (IED) Table(s)

```{r, results='asis'}
# Generate ied tables and ic50 boxplots
ic50_boxplots <- htmltools::tagList()
for (virus_to_plot in unique(data$virus)) {
  # Preprocessing
  data <- dplyr::filter(assay_df, types %in% c("x", "m"), exclude == FALSE, virus == virus_to_plot)
  model <- eval(parse(text = paste("drc::drm(", drm_model, ")")))
  # Generate plot
  suppressWarnings(ic50 <- plot_ic50_boxplot(data, model))
  ic50_boxplots[[virus_to_plot]] <- plotly::ggplotly(ic50$ic50_boxplot)
  print(knitr::kable(ic50$ied))
}
```

## IC50 Boxplot(s)

```{r}
ic50_boxplots
```

## Virus + Cell Boxplot

```{r}
# Preprocessing
boxplot_data <- dplyr::filter(boxplot_data, types %in% c("cell", "virus"))

# Generate plot
cv_boxplot <- plot_cv_boxplot(boxplot_data)
plotly::ggplotly(cv_boxplot) %>% plotly::layout(boxmode = "group")
```

# Session Info for Reproducibility

```{r}
sessionInfo()
```
