---
title: "Interactive pMN Assay Report - Generated by AutoPlate"
output:
  html_document:
    code_folding: hide
    number_sections: TRUE
params:
  drm_model: NA
---

# Input

## Load R libraries
```{r}
packages <- c("dplyr","plot.matrix","viridis","ggplot2","drc","rmarkdown","knitr","plotly","metafolio")
invisible(lapply(packages, library, character.only = TRUE))
```

## Title

```{r}
platelist_file <- "pmn_platelist.csv"
assay_df <- read.csv(platelist_file, header = TRUE, stringsAsFactors = FALSE, check.names = FALSE)
title <- paste(toString(unique(assay_df$experiment_id)), "- Bleed", toString(unique(assay_df$bleed)), "- Virus", toString(unique(assay_df$virus)))
title
```

## Let's take a look at the data

```{r, layout="l-body-outset"}
knitr::kable(head(assay_df, 3))
```

## Dimensions before exclusion

```{r}
paste0("Rows: ", dim(assay_df)[1], ", Columns: ", dim(assay_df)[2])
```

## Dimensions after exclusion

```{r}
data <- dplyr::filter(assay_df, exclude == FALSE)
paste0("Rows: ", dim(data)[1], ", Columns: ", dim(data)[2])
```

## Number of wells excluded

```{r}
dim(assay_df)[1] - dim(data)[1]
```

# Quality Control

## Average Virus + Cell Luminescence

```{r, layout="l-body-outset"}
plates <- sort(unique(assay_df$plate_number))
av_cell_lum <- lapply(plates, function(plate_n) mean(dplyr::filter(assay_df, (plate_number == plate_n) & (types == "c"))$rlu))
av_viral_lum <- lapply(plates, function(plate_n) mean(dplyr::filter(assay_df, (plate_number == plate_n) & (types == "v"))$rlu))
no_cell_wells <- lapply(plates, function(plate_n) sum(dplyr::filter(assay_df, (plate_number == plate_n))$types == "c"))
no_viral_wells <- lapply(plates, function(plate_n) sum(dplyr::filter(assay_df, (plate_number == plate_n))$types == "v"))
av_lum_df <- do.call(rbind, Map(data.frame,
  plate_number = plates,
  average_cell_luminescence = av_cell_lum,
  average_viral_luminescence = av_viral_lum,
  number_of_cell_wells = no_cell_wells,
  number_of_virus_wells = no_viral_wells
))
knitr::kable(av_lum_df)
```

## Heatmaps for The 96 Well Plate(s)

```{r}
plot_heatmap <- function(plate_number, assay_df, feature, title) {
  plate_df <- assay_df[assay_df$plate_number == plate_number, ]
  feature_list <- unlist(plate_df[[feature]], use.names = FALSE)
  vals <- matrix(feature_list, byrow = T, ncol = 12, nrow = 8)
  row.names(vals) <- LETTERS[1:8]
  # Set params for plot based on the feature
  features <- c("types", "sample_id", "dilution", "virus", "rlu", "neutralisation", "treatment", "experiment_id", "bleed", "exclude")
  fmt.cells <- c("%.5s", "%.5s", "%.5s", "%.5s", "%.0f", "%.0f", "%.5s", "%.5s", "%.5s", "%.5s")
  features <- do.call(rbind, Map(data.frame, features = features, fmt.cells = fmt.cells))
  fmt.cell <- as.character(features[feature, ]$fmt.cells)
  col <- if (feature %in% c("dilution", "rlu", "neutralisation")) viridis::viridis else rainbow
  side <- if (feature %in% c("sample_id", "treatment", "experiment_id", "bleed", "exclude")) 3 else 4
  # Generate heatmap plot
  plot(vals, col = col, fmt.cell = fmt.cell, main = paste("Plate", plate_number, title), key = list(side = side))
}

features <- c("Types", "Sample ID", "Dilution", "Virus", "RLU", "Neutralisation", "Treatment", "Experiment ID", "Bleed", "Exclude")

for (feature in features) {
  for (plate in plates) {
    plot_heatmap(plate, assay_df, gsub(" ", "_", tolower(feature), fixed = TRUE), feature)
  }
}
```

## Types boxplot

```{r}
boxplot_data <- data %>%
  dplyr::mutate(types = ifelse((types == "c"), "cell", types)) %>%
  dplyr::mutate(types = ifelse((types == "m"), "monoclonal antibody", types)) %>%
  dplyr::mutate(types = ifelse((types == "v"), "virus", types)) %>%
  dplyr::mutate(types = ifelse((types == "x"), "serum sample", types))
boxplot_data$plate_number <- as.factor(boxplot_data$plate_number)

types_boxplot <- ggplot2::ggplot(boxplot_data, ggplot2::aes(x = plate_number, y = rlu, colour = types)) +
  ggplot2::geom_boxplot() +
  ggplot2::geom_point(position=ggplot2::position_dodge(0.75)) +
  ggplot2::ylab("Raw luminescence value") +
  ggplot2::xlab("Plate number") +
  ggplot2::theme_classic() +
  ggplot2::ggtitle(title)
plotly::ggplotly(types_boxplot) %>% plotly::layout(boxmode = "group")
```

# Results

## Update order and colour of treatments

```{r}
update_cols_order <- function(data) {
  treatments <- unique(data$treatment)
  treatment_cols <- metafolio::gg_color_hue(length(treatments),0,360)
  negative_control <- c("pbs", "negative_control")
  posotive_control <- unique(data[data$type == "m",]$treatment)
  if (any(negative_control %in% tolower(treatments))) {
    negative_control_index <- which(tolower(treatments) %in% negative_control)
    treatments <- c(treatments[negative_control_index],treatments[-negative_control_index])
    treatment_cols[1] <- "grey"
  }
  if (!is.na(posotive_control)) {
    posotive_control_index <- which(treatments %in% posotive_control)
    treatments <- c(treatments[-posotive_control_index],treatments[posotive_control_index])
    treatment_cols[length(treatments)] <- "black"
  }
  data$treatment <- factor(data$treatment, levels = treatments)
  return(list(data=data, treatment_cols=treatment_cols,treatments=treatments))
}
```

## Data Exploration

```{r}
# Preprocessing
data <- dplyr::filter(data, types %in% c("x", "m"))

# Update order and colours of treatments
attach(update_cols_order(data), warn.conflicts=FALSE)

# Generate plot
data_exploration_plot <- ggplot2::ggplot(data, ggplot2::aes(x = dilution, y = neutralisation, colour = treatment)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(se = F, span = 1) +
  ggplot2::facet_wrap(. ~ virus) +
  ggplot2::ylim(c(-100, 110)) +
  ggplot2::scale_x_continuous(trans = "log10") +
  ggplot2::theme_classic() +
  ggplot2::scale_colour_manual(breaks=treatments,values=treatment_cols) +
  ggplot2::ylab("Neutralisation") +
  ggplot2::xlab("Dilution") +
  ggplot2::ggtitle(title)
plotly::ggplotly(data_exploration_plot)
```

## DRM model

```{r}
params$drm_model
```

## Dose Response Curve(s)

```{r}
plot_drc <- function(data,virus_to_plot,drm_model) {
  # Filter to keep only one virus
  data <- dplyr::filter(data, virus == virus_to_plot)

  # Update order and colours of treatments
  attach(update_cols_order(data), warn.conflicts=FALSE)

  # Preprocessing
  data$sample_id <- unlist(data$sample_id)
  model <- eval(parse(text = paste("drc::drm(", drm_model, ")")))
  n <- 100
  new_dilution <- exp(seq(log(min(data$dilution)), log(max(data$dilution)), length.out = n))
  sample_ids <- unique(data$sample_id)
  new_data <- expand.grid(new_dilution, sample_ids)
  names(new_data) <- c("dilution", "sample_id")
  new_data$treatment <- data$treatment[match(new_data$sample_id, data$sample_id)]
  new_data$neutralisation <- predict(model, newdata = new_data, )

  # Generate plot
  drc_plot <- ggplot2::ggplot(new_data, ggplot2::aes(x = dilution, y = neutralisation, colour = treatment, group = sample_id)) +
    ggplot2::geom_line() +
    ggplot2::geom_point(data = data, ggplot2::aes(y = neutralisation)) +
    ggplot2::facet_wrap(.~treatment) +
    ggplot2::scale_x_continuous(trans = "log10") +
    ggplot2::theme_classic() +
    ggplot2::scale_colour_manual(breaks=treatments,values=treatment_cols) +
    ggplot2::theme(strip.background = ggplot2::element_blank()) +
    ggplot2::ylab("Neutralisation") +
    ggplot2::xlab("Dilution") +
    ggplot2::ggtitle(paste(unique(data$experiment_id), "- Bleed", unique(data$bleed), "- Virus", unique(data$virus)))
  # plotly::ggplotly(drc_plot)
  return(drc_plot)
}

# Produce a seperate plot for each virus
htmltools::tagList(lapply(unique(data$virus), function(virus_to_plot) { 
  plotly::ggplotly(plot_drc(data,virus_to_plot,params$drm_model)) 
}))
```

## Individual Effective Dose (IED) Table(s)

```{r, results='asis'}
plot_ic50 <- function(data,virus_to_plot,drm_model) {
  # Filter to keep only one virus
  data <- dplyr::filter(data, virus == virus_to_plot)

  # Update order and colours of treatments
  attach(update_cols_order(data), warn.conflicts=FALSE)

  # Preprocessing
  data$sample_id <- as.character(data$sample_id)
  model <- eval(parse(text = paste("drc::drm(", drm_model, ")")))
  ied <- as.data.frame(drc::ED(model, 50, display = FALSE))
  ied$sample_id <- gsub("e:|:50", "", row.names(ied))
  ied$treatment <- data$treatment[match(ied$sample_id, data$sample_id)]
  ied$plate_number <- data$plate_number[match(ied$sample_id, data$sample_id)]
  ied$virus <- data$virus[match(ied$sample_id, data$sample_id)]
  control_median <- median(ied[tolower(ied$treatment) %in% tolower(c("PBS", "negative_control")),]$Estimate)
  # Average Neutralisation
  avied <- dplyr::summarise(dplyr::group_by(ied, treatment), av = median(Estimate))
  ied_order <- avied$treatment[order(avied$av)]

  # Generate plot
  ic50_boxplot <- ggplot2::ggplot(ied, ggplot2::aes(x = treatment, y = Estimate, colour = treatment)) +
    ggplot2::geom_boxplot() +
    ggplot2::geom_point() +
    ggplot2::scale_x_discrete(limits = ied_order) +
    ggplot2::ylab("Individual IC50 log10") +
    ggplot2::xlab("Treatment") +
    ggplot2::theme_classic() +
    ggplot2::scale_colour_manual(breaks=treatments,values=treatment_cols) +
    ggplot2::ggtitle(paste(unique(data$experiment_id), "- Bleed", unique(data$bleed), "- Virus", unique(data$virus))) +
    ggplot2::coord_flip() +
    ggplot2::geom_hline(yintercept=c(control_median), linetype="dotted", color="grey")
  return(list(ic50_boxplot=ic50_boxplot, ied=ied))
}

# Generate ied tables and ic50 boxplots
ic50_boxplots <- htmltools::tagList()
for (virus_to_plot in unique(data$virus)) {
  suppressWarnings(ic50_boxplot <- plot_ic50(data,virus_to_plot,params$drm_model))
  ic50_boxplots[[virus_to_plot]] <- plotly::ggplotly(ic50_boxplot$ic50_boxplot)
  print(knitr::kable(ic50_boxplot$ied))
}
```

## IC50 Boxplot(s)

```{r}
ic50_boxplots
```

## Virus + Cell Boxplot

```{r}
boxplot_data <- dplyr::filter(boxplot_data, types %in% c("cell", "virus"))

cv_boxplot <- ggplot2::ggplot(boxplot_data, ggplot2::aes(x = types, y = rlu, colour = plate_number)) +
  ggplot2::geom_boxplot() +
  ggplot2::geom_point(position = ggplot2::position_dodge(0.75)) +
  ggplot2::scale_y_continuous(trans = "log10") +
  ggplot2::ylab("Log10 raw luminescence value") +
  ggplot2::xlab("Cell only or Virus only") +
  ggplot2::theme_classic() +
  ggplot2::ggtitle(title)
plotly::ggplotly(cv_boxplot) %>% plotly::layout(boxmode = "group")
```

# Session Info for Reproducibility

```{r}
sessionInfo()
```
